---
title: "RAMP Upwelling Dashboard"
date: 09-19-2025
date-modified: last-modified
format:
  dashboard:
    orientation: columns
---

Last updated: **{{< meta date-modified >}}**

This is an example dashboard demonstrating how to acquire, summarize, and visualize environmental data from a [product on ERDDAP](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdQCwindproducts3day). Below are some examples of ways that these results can be visualized through tables and plots.

<!-- Specify card headers as bold -->
<style>
.card-header {
  font-weight: bold;
}
</style>


## Column {width=30%}

```{r}
#| label: tbl

library(tidyverse)
library(DT)
library(plotly)

# Load data
df <- read_csv("ERDDAP_example/Monthly RAMP upwelling.csv")

df |> 
  mutate(ekman_upwelling = round(ekman_upwelling, 8)) |> 
  arrange(desc(time)) |> 
  datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           bPaginate = TRUE, # Enable pagination
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


## Column {width=60%}

```{r}
#| content: card-toolbar

library(bslib)
library(bsicons)

tooltip(
  bs_icon("info-circle"),
  "Mean monthly Ekman upwelling per RAMP zone is shown, where data was acquired from an ERDDAP product. The dataset begins in January 2024.",
  placement = "left"
)
```

```{r}
#| title: Upwelling over time
#| label: ts-plot

ggplotly(
  ggplot(df) +
    geom_hline(yintercept = 0, linewidth = 0.5, color = "grey80") +
    geom_line(aes(x = time, y = ekman_upwelling, colour = FishingZone), linewidth = 1) +
    scale_color_brewer("RAMP Zone", palette = "Set2") +
    labs(x = "Month", y = "Ekman Upwelling (m s-1)") +
    theme_bw(base_size = 16) +
    theme(panel.grid = element_blank())
)
```




```{r}
#| content: card-toolbar

tooltip(
  bs_icon("info-circle"),
  "This plot shows the monthly mean Ekman upwelling values per RAMP zone for the last 9 months. RAMP zones are numbered from North to South (1 to 6). RAMP Zone 7 has be omitted from these maps for easier viewing.",
  placement = "left"
)
```

```{r}
#| title: Upwelling over time and space
#| label: map

library(sf)

ramp <- st_read("ERDDAP_example/Monthly_RAMP_upwelling.geojson", quiet = TRUE)

last_9_months <- ramp |> 
  distinct(time) |> 
  arrange(desc(time)) |> 
  slice(1:9) |> 
  pull(time)

ramp |> 
  filter(time %in% last_9_months,
         FishingZone != "Zone 7") |> 
  ggplot(aes(group = FishingZone)) +
  geom_sf(aes(fill = ekman_upwelling), alpha = 0.7, color = "black") +
  scale_fill_viridis_c("Ekman\nUpwelling\n(m s-1)") +
  geom_sf_text(data = ramp |> 
                 filter(time == last_9_months[9],
                        FishingZone != "Zone 7"),
               aes(label = FishingZone)) +
  guides(fill = guide_colorbar(barheight = 20, barwidth = 1)) +
  theme_void(base_size = 16) +
  theme(
    legend.position = "right",
    legend.margin = margin(t = 0, r = 10, b = 0, l = 20, unit = "pt") # Adjust as needed
  ) +
  facet_wrap(~ time)

```



## Column {width=10%}

::: {.valuebox icon="calendar-event" color="blue"}
Latest data

`r last(df$time)`
:::

::: {.valuebox icon="tsunami" color="warning"}
Latest min upwelling

`r df |> filter(time == last(df$time)) |> pull(ekman_upwelling) |> min(na.rm = TRUE)`
:::

::: {.valuebox icon="tsunami" color="danger"}
Latest min upwelling

`r df |> filter(time == last(df$time)) |> pull(ekman_upwelling) |> max(na.rm = TRUE)`
:::